---
title: "The backtest and its intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The backtest and its intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

```{r}
library(foresightr)
data <- read.csv(system.file("extdata", "piaui_revenue.csv", package = "foresightr"),
                 comment.char = "#")
fpe <- ts(data$fpe / 1e6, start = c(2017, 3), frequency = 12)  # BRL million
```

## Replaying the past

A rolling-origin backtest stands at each of the last `origins` periods of the
series, fits every candidate on the data before that point only, and
forecasts up to `horizon` periods ahead. Each forecast is compared with what
actually happened. Nothing a model sees at an origin comes from after it.

```{r}
bt <- backtest(fpe, origins = 36, horizon = 12)
bt$origins
bt$first_origin   # position of the first period forecast
```

Series shorter than `min_train` plus the origins get fewer origins; `window`
trains on the last observations only, for series whose behaviour changed.

## The error by horizon

Errors grow with the horizon. Each candidate keeps them one horizon at a time:
the pairs evaluated, the mean absolute percentage error, the bias (positive
when forecasts ran high), MAE, RMSE and MASE.

```{r}
acc <- bt$candidates[[bt$best]]$accuracy
acc[c(1, 3, 6, 12), ]
```

The ranking averages the chosen `metric` over the horizons. The last
candidate is the simple average of the best `combine` models, often better
than either:

```{r}
bt$ranking[order(bt$ranking$score), c("name", "score")][1:5, ]
```

Changing the metric changes the ranking only when models trade accuracy in
different ways; `"mase"` scales the errors by the seasonal naive error of each
training set, which makes series of different sizes comparable.

## Intervals from the errors actually made

An interval here is not derived from a distribution. At each horizon, the
backtest collects the relative errors `actual / forecast - 1` of the
candidate and takes their quantiles: the 80% interval goes from the 10% to the
90% quantile. The bands are kept in relative terms and applied to the final
forecast.

```{r}
bt$candidates[[bt$best]]$bands[c(1, 6, 12), ]
bt$forecast[c(1, 6, 12), ]
```

With 36 origins, the 95% bounds rest on few errors at long horizons; they are
honest about the past but not smooth. More origins give steadier bands, at the
cost of judging models on older data.

The trajectories of the backtest are kept, one row per origin, should you want
another summary of them:

```{r}
dim(bt$candidates[[bt$best]]$trajectories)
```

## Totals

Budgets are about totals: the revenue of the rest of the year, not of each
month. The backtest also measures the error of the sum of the first k periods,
so the total gets its own interval:

```{r}
total_forecast(bt, 6)
```

Adding up the six monthly bounds would overstate the uncertainty: monthly
errors partly cancel in a sum.

```{r}
colSums(bt$forecast[1:6, c("mean", "lower_80", "upper_80")])
```

## Looking at it

```{r}
plot(bt)
```

Any candidate can be plotted or inspected by name:

```{r}
plot(bt, candidate = "seasonal_naive")
```
