---
title: "Decomposition and cleaning"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decomposition and cleaning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

```{r}
library(foresightr)
```

## STL

`decompose_stl()` splits a series into trend, seasonal pattern and remainder
by LOESS. Its defaults are those of `stats::stl()`, and so are its numbers.

```{r}
d <- decompose_stl(log(AirPassengers), seasonal_window = 13)
d
plot(d)
```

The strengths, from 0 to 1, say how much of the variation the trend and the
seasonal pattern explain; `seasonal_strength()` computes the latter directly.
`robust = TRUE` down-weights outliers.

## Several seasonal periods

Daily data often has a weekly and a monthly pattern. `decompose_mstl()`
estimates each in turn:

```{r}
t <- 0:419
weekly <- c(5, 0, -2, -3, 0, 1, -1)
daily <- 100 + 0.05 * t + weekly[t %% 7 + 1] + 8 * sin(2 * pi * t / 30) +
  2 * sin(t * 1.7) * cos(t * 0.3)
d <- decompose_mstl(daily, periods = c(7, 30))
d$seasonal_strength
plot(d)
```

The same decomposition serves to forecast: `model_decomposed()` runs any model
on the seasonally adjusted series and adds the patterns back.

```{r}
f <- forecast_model(model_decomposed(model_drift(), periods = c(7, 30)), daily, 14, period = 7)
round(f, 1)
```

## Gaps and outliers

```{r}
y <- log(AirPassengers)
y[30] <- y[30] + 0.8     # a typing error
y[100] <- y[100] - 0.7   # a strike
y[61:62] <- NA           # months never recorded
find_outliers(y)
```

`fill_gaps()` fills the missing values following the seasonal pattern;
`clean_series()` also replaces the outliers:

```{r}
cleaned <- clean_series(y)
round(exp(cbind(dirty = y, clean = cleaned)[c(30, 61, 62, 100), ]))
```

Models require complete series, so cleaning comes first when data has gaps.
Whether an outlier should be replaced is a judgement about the data: a strike
really happened, and removing it may hide a risk that could happen again.
