---
title: "Regressors and events"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regressors and events}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

```{r}
library(foresightr)
path <- system.file("extdata", "piaui_revenue.csv", package = "foresightr")
data <- read.csv(path, comment.char = "#")
icms <- ts(data$icms / 1e6, start = c(2017, 3), frequency = 12)
```

External information enters a model in three ways: as regressors of an
ARIMA model, as a price index that deflates a log-linear regression, and as
dated events of a Prophet-style model. Regressors and events must also cover
the periods to be forecast; asking for a forecast beyond them is an error.

## Regression with ARIMA errors

Tax revenue follows prices. With the log of a price index as a regressor, the
errors of the regression follow an ARIMA model, and both are estimated
together. Here the model is fitted on the first 100 months, so that the index
covers the 12 months forecast.

```{r}
model <- model_arima(c(0, 1, 1), c(0, 1, 1),
                     regressors = data.frame(ipca = log(data$ipca_index)))
fit <- fit_model(model, log(window(icms, end = time(icms)[100])))
fit$details$regression
exp(predict(fit, 6))
```

The same `regressors` argument is taken by `model_arima()` and
`model_auto_arima()`, as a data frame, a matrix or a named list with one row
per period from the first observation on.

## Fourier terms

Sine and cosine pairs describe a seasonal pattern with few parameters, and the
period need not be a whole number. `fourier_terms()` gives them for as many
rows as needed:

```{r}
y <- log(AirPassengers)
x <- fourier_terms(12, 3, length(y) + 12)
fit <- fit_model(model_arima(c(1, 1, 1), constant = TRUE, regressors = x), y)
names(fit$details$regression)
round(exp(predict(fit, 12)))
```

`seasonal_dummies()` gives one dummy per season instead; combine sets of
regressors with `cbind()`.

## A deflated regression

`model_log_linear(deflator = )` fits trend and seasonality on the deflated
series. The future of the index is not needed: the forecasts are inflated
back at the growth of the index over the last twelve months.

```{r}
deflated <- model_log_linear(deflator = data$ipca_index)
fit <- fit_model(deflated, icms)
fit$params[c("trend_growth_pct", "inflation_pct")]
predict(fit, 6)
```

## Events and steps

A Prophet-style model takes events, which happen at given positions (counted
from 1 at the first observation, future ones included), and steps, lasting
changes of level from a position on. Their effects are estimated with the
trend and seasonality.

```{r}
set.seed(1)
t <- 1:120
sales <- 200 + 0.8 * t + 5 * sin(2 * pi * t / 12) + rnorm(120, sd = 2)
campaigns <- c(11, 35, 59, 83, 107)
sales[campaigns] <- sales[campaigns] + 20     # a campaign every other November
sales[t >= 81] <- sales[t >= 81] - 15          # a new law cuts the level
sales <- ts(sales, frequency = 12)

model <- model_prophet(changepoints = 0,
                       events = list(campaign = c(campaigns, 131)),
                       steps = list(new_law = 81))
fit <- fit_model(model, sales)
round(fit$details$effects, 1)
```

The effects come out close to the +20 and -15 put in, and the forecast
includes the campaign planned for position 131.
